---
title: "Are Songs Getting faster? A study of top songs over the last 24 years"
author: "Tim Stephens"
date: "01/02/2021"
output: html_document
---
#Background

This short investigation hopes to determine if top songs are getting faster. Top songs from 1996-2020 (the years I have been alive) were analyzed to determine if there is a trend with higher tempo songs gaining more traction. Top songs were taken from the "Billboard Magazine's Top Hot 100" from each year. It was found that there was no relationship between tempo of songs and time - successful songs vary in tempo year-to-year. 

```{r, echo = FALSE, message = FALSE}

###Workspace Set-up###

library(devtools)
library(here)
library(usethis)
library(spotifyr)
library(tidyverse)

#add your spotify clientID, secret
usethis::edit_r_environ() 

#function to pull trackname, valence, tempo, and album release year from the dataset. 
get_artist_trackName_valence_tempo_year <- 
  function(artist, track){
    artist %>% 
    select(artist_name, track_name, valence, tempo, album_release_year) %>% 
    filter(track_name == track) %>% 
    filter(row_number() == 1)
  }


#all years were analyzed the same. Just 1996 is commented. 

#1996
#get the features from the top artist of that year: 
top_1996_los_del_rio <- get_artist_audio_features('los_del_rio')

#save the information in a local directory:
saveRDS(top_1996_los_del_rio, here("MusicData_noPush/los_del_rio.rds"))

#load in from local: 
los_del_rio <- readRDS(here("MusicData_noPush/los_del_rio.rds"))

#call the function, get the necessary info from the dataset
top_1996 <- get_artist_trackName_valence_tempo_year(los_del_rio, "Macarena")

#show the results
top_1996



#1997
top_1997_elton_john <- get_artist_audio_features('elton_john')
saveRDS(top_1997_elton_john, here("MusicData_noPush/elton_john.rds"))
elton <- readRDS(here("MusicData_noPush/elton_john.rds"))
top_1997 <- get_artist_trackName_valence_tempo_year(elton, "Candle In The Wind")
top_1997


#1998
top_1998_next <- get_artist_audio_features('next')
saveRDS(top_1998_next, here("MusicData_noPush/next.rds"))
next_the_artist <- readRDS(here("MusicData_noPush/next.rds"))
top_1998 <- get_artist_trackName_valence_tempo_year(next_the_artist, "Too Close")
top_1998


#1999
top_1999_cher <- get_artist_audio_features('cher')
saveRDS(top_1999_cher, here("MusicData_noPush/cher.rds"))
cher <- readRDS(here("MusicData_noPush/cher.rds"))
top_1999 <- get_artist_trackName_valence_tempo_year(cher, "Believe")
top_1999

#2000
top_2000_faith_hill <- get_artist_audio_features('faith_hill')
saveRDS(top_2000_faith_hill, here("MusicData_noPush/faith_hill.rds"))
faith_hill <- readRDS(here("MusicData_noPush/faith_hill.rds"))
top_2000 <- get_artist_trackName_valence_tempo_year(faith_hill, "Breathe")
top_2000

#2001
top_2001_lifehouse <- get_artist_audio_features("lifehouse")
saveRDS(top_2001_lifehouse, here("MusicData_noPush/lifehouse.rds"))
lifehouse <- readRDS(here("MusicData_noPush/lifehouse.rds"))
top_2001 <- get_artist_trackName_valence_tempo_year(lifehouse, "Hanging By A Moment")
top_2001

#2002
top_2002_nickelback <- get_artist_audio_features("nickelback")
saveRDS(top_2002_nickelback, here("MusicData_noPush/nickelback.rds"))
nickelback <- readRDS(here("MusicData_noPush/nickelback.rds"))
top_2002 <- get_artist_trackName_valence_tempo_year(nickelback, "How You Remind Me")
top_2002

#2003
top_2003_50c <- get_artist_audio_features("50_cent")
saveRDS(top_2003_50c, here("MusicData_noPush/fiddy.rds"))
fiddy <- readRDS(here("MusicData_noPush/fiddy.rds"))
top_2003 <- get_artist_trackName_valence_tempo_year(fiddy, "In Da Club")
top_2003

# #2004 - Oddly, Usher's masterpiece: Yeah! Ft. lil Jon and Ludacris, is not in the dataset.
# top_2004_usher <- get_artist_audio_features("Usher")
# saveRDS(top_2004_usher, here("usher.rds"))
# usher <- readRDS(here("usher.rds"))
# top_2004 <- get_artist_trackName_valence_tempo(usher, "Yeah! feat. Lil Jon & Ludacris")
# top_2004

#2005
top_2005_mariah <- get_artist_audio_features("mariah_carey")
saveRDS(top_2005_mariah, here("MusicData_noPush/mariah.rds"))
mariah <- readRDS(here("MusicData_noPush/mariah.rds"))
top_2005 <- get_artist_trackName_valence_tempo_year(mariah, "We Belong Together")
top_2005


#2006
top_2006_daniel <- get_artist_audio_features("daniel powter")
saveRDS(top_2006_daniel, here("MusicData_noPush/daniel.rds"))
daniel <- readRDS(here("MusicData_noPush/daniel.rds"))
top_2006 <- get_artist_trackName_valence_tempo_year(daniel, "Bad Day")
top_2006


#2007
top_2007_bey <- get_artist_audio_features("Beyoncé")
saveRDS(top_2007_bey, here("MusicData_noPush/Beyoncé.rds"))
beyonce <- readRDS(here("MusicData_noPush/Beyoncé.rds"))
top_2007 <- get_artist_trackName_valence_tempo_year(beyonce, "Irreplaceable")
top_2007


#2008 
top_2008_flo <- get_artist_audio_features("flo-rida")
saveRDS(top_2008_flo, here("MusicData_noPush/flo.rds"))
flo <- readRDS(here("MusicData_noPush/flo.rds"))
top_2008 <- get_artist_trackName_valence_tempo_year(flo, "Low (feat. T-Pain)")
top_2008 

#2009
top_2009_peas <- get_artist_audio_features("black eyed peas")
saveRDS(top_2009_peas, here("MusicData_noPush/peas.rds"))
peas <- readRDS(here("MusicData_noPush/peas.rds"))
top_2009 <- get_artist_trackName_valence_tempo_year(peas, "Boom Boom Pow")
top_2009 


#2010 - No tik tok either!
top_2010_kesha <- get_artist_audio_features("ke$ha")
saveRDS(top_2010_kesha, here("MusicData_noPush/kesha.rds"))
kesha <- readRDS(here("MusicData_noPush/kesha.rds"))
top_2010 <- get_artist_trackName_valence_tempo_year(kesha, "Tik Tok")
top_2010 

#2011
top_2011_adele <- get_artist_audio_features("adele")
saveRDS(top_2011_adele, here("MusicData_noPush/adele.rds"))
adele <- readRDS(here("MusicData_noPush/adele.rds"))
top_2011 <- get_artist_trackName_valence_tempo_year(adele, "Rolling in the Deep")
top_2011

#2012
top_2012_gotye <- get_artist_audio_features("Gotye")
saveRDS(top_2012_gotye, here("MusicData_noPush/gotye.rds"))
gotye <- readRDS(here("MusicData_noPush/gotye.rds"))
top_2012 <- get_artist_trackName_valence_tempo_year(gotye, "Somebody That I Used To Know")
top_2012


#2013
top_2013_mack <- get_artist_audio_features("macklemore")
saveRDS(top_2013_mack, here("MusicData_noPush/mack.rds"))
mack <- readRDS(here("MusicData_noPush/mack.rds"))
top_2013 <- get_artist_trackName_valence_tempo_year(mack, "Thrift Shop (feat. Wanz)")
top_2013

#2014
top_2014_pharrell <- get_artist_audio_features("Pharrell")
saveRDS(top_2014_pharrell, here("MusicData_noPush/ph.rds"))
ph <- readRDS(here("MusicData_noPush/ph.rds"))
top_2014 <- get_artist_trackName_valence_tempo_year(ph, 'Happy - From "Despicable Me 2"')
top_2014


#2015
top_2015_mark <- get_artist_audio_features("Mark Ronson")
saveRDS(top_2015_mark, here("MusicData_noPush/mark.rds"))
mark <- readRDS(here("MusicData_noPush/mark.rds"))
top_2015 <- get_artist_trackName_valence_tempo_year(mark, "Uptown Funk (feat. Bruno Mars)")
top_2015


#2016 no Love Yourself either
top_2016_bieber <- get_artist_audio_features("Justin Bieber")
saveRDS(top_2016_bieber, here("MusicData_noPush/biebs.rds"))
biebs <- readRDS(here("MusicData_noPush/biebs.rds"))
top_2016 <- get_artist_trackName_valence_tempo_year(biebs, "Love Yourself")
top_2016

#2017 - No sheeran...weird 
top_2017_sheeran <- get_artist_audio_features("Ed Sheeran")
saveRDS(top_2017_sheeran, here("MusicData_noPush/sheeran.rds"))
sheeran <- readRDS(here("MusicData_noPush/sheeran.rds"))
top_2017 <- get_artist_trackName_valence_tempo_year(sheeran, "Shape of You")
top_2017

#2018
top_2018_Drake <- get_artist_audio_features("Drake")
saveRDS(top_2018_Drake, here("MusicData_noPush/drake.rds"))
drake <- readRDS(here("MusicData_noPush/drake.rds"))
top_2018 <- get_artist_trackName_valence_tempo_year(drake, "God's Plan")
top_2018

#2019 #couldnt find it (old town Road)
#top_2019_nas <- get_artist_audio_features("lil nas x")



#2020
top_2020_weeknd <- get_artist_audio_features("The Weeknd")
saveRDS(top_2020_weeknd, here("MusicData_noPush/weeknd.rds"))
weeknd <- readRDS(here("MusicData_noPush/weeknd.rds"))
top_2020 <- get_artist_trackName_valence_tempo_year(weeknd, "Blinding Lights")
top_2020

```

```{r, echo = FALSE, message = FALSE}

#combine all individual tables into one large table using rbind

raw_data <- 
  rbind(top_1996, 
        top_1997, 
        top_1998, 
        top_1999, 
        top_2000, 
        top_2001, 
        top_2002,
        top_2003,
        #top_2004,
        top_2005,
        top_2006,
        top_2007,
        top_2008,
        top_2009,
        #top_2010,
        top_2011,
        top_2012,
        top_2013,
        top_2014,
        top_2015,
        #top_2016,
        #top_2017,
        top_2018,
        #top_2019,
        top_2020
        )


```

```{r, echo = FALSE}
#convert table to tibble with better names
music_data <- 
  tibble(artist = raw_data$artist_name,
         song = raw_data$track_name,
         valence = raw_data$valence,
         temp = raw_data$tempo,
         year = raw_data$album_release_year)
  
#create a scatter plot to compare tempo over time
music_chart_tempo <-
  music_data %>% 
  ggplot(mapping = aes(x = year, y = temp))+
  geom_point(colour = "darkgreen") + #green b/c spotify
  theme_minimal() + 
  labs(x = "Year",
       y = "Tempo (beats-per-minute)",
       title = "There is No Clear Relationship Between Time and Tempo of Top Songs",
       subtitle = "Songs are from the past 24 years",
       caption = "Data from Spotify")

music_chart_tempo

#save the file
ggsave(here("outputs/Tempo_Vs_Time.pdf"), width = 40, height = 20, units = "cm")

```



